manuscript_outline

### 2025-07-10 docs/manuscript/paper.Rmd
new table 2 ref
removed taxa with 2 or less temp data points from table 2

@@

|:--------------|:--------------|:--------------|:--------------|:--------------|

@@

|              | Lepidoptera   | *Choristoneura* | *C. fumiferana*    | Régnière et al. (2012) [10.1016/j.jinsphys.2012.01.010](https://doi.org/10.1016/j.jinsphys.2012.01.010) |

@@

| Copepoda     | Calanoida     | 13 genera       |                    | Forster et al. (2011) <https://doi.org/10.1086/662174>                                                  |

@@

@@

```{r tail-frog-pheno}
# Read data
ascaphus_siegel_data <- readRDS(here("data/ascaphus_siegel_data.rds")) |>
  mutate(date = ymd(tim.date))

@@

The dataset included `r nrow(model_data)` observations of predicted days to hatching, drawn from `r length(unique(model_data$comid))` stream reaches (labeled with Common Identifiers, i.e., COMIDs) nested within four mountain ranges: Olympic National Park, the Cascade Mountains, the Siskiyou Mountains, and the Willapa Hills. Each COMID had between 20 and 30 years of predictions.

@@

```{r frog-mods}
# mod <- lm(days ~ year * comid + Mountain, data = model_data)

@@

This model explained \~`r round(perf[1,4], 2)*100`% of the variation in developmental timing (adjusted R² = `r round(perf[1,4], 3)`), with year having a significant weak and negative effect (`r round(params[2,2], 2)` days/year, p = 0.001), indicating an overall trend toward faster development consistent with regional warming (Figure \@ref(fig:frog-preds)A). COMID effects captured the persistent differences in thermal regimes among streams, with estimated marginal means ranging from \~15 to \~44 days to hatch across sites (Figure \@ref(fig:frog-preds)B).

@@

```{r frog-preds, fig.height=3 , fig.width=8, fig.align='center', fig.cap="Modeled changes in developmental timing for coastal tailed frogs (Ascaphus truei) from 1991 to 2020. Panel A: Marginal mean trend across all sites showing a significant decline in days to hatching over time. Panel B: Site-specific (COMID) trends in predicted days to hatch, colored by mountain range. Estimates are based on the best fitting model, with stream temperatures from @siegel2023 and developmental timing modeled using hatchR. Shaded ribbons represent 95% confidence intervals."}

>>>>

+--------------+---------------+-----------------+--------------------+---------------------------------------------------------------------------------------------------------+

@@

+:=============+:==============+:================+:===================+:========================================================================================================+

@@

+--------------+---------------+-----------------+--------------------+---------------------------------------------------------------------------------------------------------+
|              | Lepidoptera   | *Choristoneura* | *C. occidentalis*  | Régnière et al. (2012) [10.1016/j.jinsphys.2012.01.010](https://doi.org/10.1016/j.jinsphys.2012.01.010) |
+--------------+---------------+-----------------+--------------------+---------------------------------------------------------------------------------------------------------+

@@

+--------------+---------------+-----------------+--------------------+---------------------------------------------------------------------------------------------------------+
| Copepoda     | Calanoida     | 11 genera       |                    | Forster et al. (2011) <https://doi.org/10.1086/662174>                                                  |
+--------------+---------------+-----------------+--------------------+---------------------------------------------------------------------------------------------------------+

@@

+--------------+---------------+-----------------+--------------------+---------------------------------------------------------------------------------------------------------+
| Asteroidea   | Valvatida     | 3 genera        |                    | Hoegh-Guldberg & Pearse (1995)                                                                          |
|              |               |                 |                    |                                                                                                         |
|              |               |                 |                    | <https://sci-hub.st/https://doi.org/10.1093/icb/35.4.415>                                               |
+--------------+---------------+-----------------+--------------------+---------------------------------------------------------------------------------------------------------+

@@

```{r tail-frog-pheno, eval = FALSE}
# Read data
ascaphus_siegel_data <- readRDS(here("data/ascaphus_siegel_data.rds")) |>
  mutate(date = ymd(tim.date))

@@

The dataset included r nrow(model_data) observations of predicted days to hatching, drawn from r length(unique(model_data\$comid)) stream reaches (labeled with Common Identifiers, i.e., COMIDs) nested within four mountain ranges: Olympic National Park, the Cascade Mountains, the Siskiyou Mountains, and the Willapa Hills. Each COMID had between 20 and 30 years of predictions.

@@

```{r frog-mods, eval = FALSE}
# mod <- lm(days ~ year * comid + Mountain, data = model_data)

@@

This model explained r round(perf[1,4], 2)\*100 % of the variation in developmental timing (adjusted R² = r round(perf[1,4], 3)), with year having a significant weak and negative effect (r round(params[2,2], 2) days/year, p = 0.001), indicating an overall trend toward faster development consistent with regional warming (Figure \@ref(fig:frog-preds)A). COMID effects captured the persistent differences in thermal regimes among streams, with estimated marginal means ranging from \~15 to \~44 days to hatch across sites (Figure \@ref(fig:frog-preds)B).

@@

```{r frog-preds, fig.height=3 , fig.width=8, fig.align='center', fig.cap="Modeled changes in developmental timing for coastal tailed frogs (Ascaphus truei) from 1991 to 2020. Panel A: Marginal mean trend across all sites showing a significant decline in days to hatching over time. Panel B: Site-specific (COMID) trends in predicted days to hatch, colored by mountain range. Estimates are based on the best fitting model, with stream temperatures from @siegel2023 and developmental timing modeled using hatchR. Shaded ribbons represent 95% confidence intervals.", eval = FALSE}
